---
title: "Central Limit Theorem"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='CLT/figs/')
```


# Exploration of the Central Limit Theorem
The Central Limit Thereom is one of the most fundamental concepts in statistics. In this document, I will be visualizing the central limit theorem to better drive my understand of this concept. The theorem itself revolves around the sampling process, in which we try to approximate the distribution of the population by drawing values from the whole population. The underlying assumption that we cannot sample the whole population, so we sample until we feel that we have an accurate estimation of our population. 


For this example, we will assume our population is 100,000 males that live in a hypothetical country. We can say that the height of all the males ranges from 1.5 m (~4.9 feet) to 2 m (~6.6 feet). Our population follows the following distribution (mean is shown by a vertical red line):
```{r Height_Distribution}
height_dist <- runif(100000, 1.5, 2)
mean_height <- mean(height_dist)
hist(height_dist, 
     main = 'Height Distribution for 100,000 Males',
     xlab = 'Height in M',
     breaks = 100)
abline(v=mean_height,col="red")
```

Given this population, we hope to run an experiment to try and approximate this distribution. Since we "can't" sample all 100000 males, we decide our best course of action is to complete simple random sampling (SRS). Below is a function to create samples with size N for reps repetitions from a distribution dist: 
```{r Sampl_Function}
# Function takes the sample size N, the number of sampling repetitions to complete and distribution to sample from

sampl <- function(n, reps, dist) {
  samples <- replicate(reps, sample(dist, n))
  sample_means <- colMeans(samples)
  return(sample_means)
}
```

Lets start with samples of size 3 for 100 repetitions from our testing distribution:

```{r Sample_Size_3_1000_reps}
means <- sampl(3, 1000, height_dist)
hist(means)
```

This is where the central limit theorem comes into play: as we increase our sample size (N), the resulting sampling distribution (the distribution of our sample means) will begin to resemble the normal distribution:

```{r Sample_Size_10_1000_reps}
hist(sampl(10, 1000, height_dist))
```

```{r Sample_Size_20_1000_reps}
hist(sampl(20, 1000, height_dist))
```

```{r Sample_Size_50_1000_reps}
hist(sampl(50, 1000, height_dist))
```


Through the proggression of the above distributions, we can see that by only changing the sample size N, we have approximated a normal distribution.